---
output: pdf_document
---
```{r}
library(netmeta)
```

# Function to generate two-arm data
$n$ denotes number of treatments and $m$ denotes number of pairwise comparisons

Design matrix is denoted by $X\in\mathbb{R}^{m\times (n-1)}$, $X_{ij}=1$ indicating in comparison $i$, treatment $j$ is compared to the reference treatment.

The true effect difference is denoted by $\mu\in\mathbb{R}^{n-1}$, $\mu_j$ denotes the effect difference between treatment $j$ and reference treatment

Data is simulated using the following formula:
$$\theta_i\sim N(Xd, \phi s_i^2 + \tau^2)$$
$\phi\neq0$

Fixed-effect model: $\phi=1,\tau=0$

Additive effect: $\phi=1, tau\neq 0$

Multiplicative effect: $\phi\neq 1, \tau=0$

```{r}
# X: m*n-1 design matrix
# se: m-dimensional vector, standard errors
# d: (n-1)-dimensional vector, true effect sizes
# tau: scalar, additive parameter, can also handle m-dim vector values
# phi: scalar, multiplicative parameter
# ref: name of the reference treatment

# generate data that can be passed to netmeta() directly
# assume a fixed-effect model by default (tau = 0, phi = 1)
# output  treat1, treat2, studlab, seTE, TE

generate_2arm <- function(X, d, se, phi = 1, tau = 0, ref = "ref") {
  
  m <- nrow(X)   # number of pairwise comparisons
  n <- ncol(X)   # number of treatments
  
  treatments <- colnames(X)   # treatment names
  # apply(X,1,fun) apply fun to each row of X
  trt1 <- apply(X, 1, function(row) {
    treatments[row == 1]
  })
  trt2 <- rep(ref, m)
  nmadata <- as.data.frame(cbind(trt1, trt2))
  colnames(nmadata) <- c("treat1", "treat2")
  
  # study labels
  nmadata$studlab <- paste0("Study", 1:m)
  
  # standard errors 
  nmadata$seTE <- se
  
  # effect sizes
  theta <- X %*% d
  TE <- rnorm(m, mean = theta, sd = sqrt(phi * se^2 + tau^2))
  nmadata$TE <- TE
  
  return(nmadata)
}
```

Example
```{r}
set.seed(123)
X <- matrix(c(rep(c(1,0),3), rep(c(0,1),3)), nrow =6, byrow = TRUE)
colnames(X) <- c("T1", "T2")
d <- c(5,10)
se <- rep(1,6)
nmadata<-generate_2arm(X, d, se)
nmadata
net <- netmeta(TE, seTE, treat1, treat2, studlab, data = nmadata)
net$TE.common
```

quick simulation to check if unbiased
```{r eval=FALSE, include=FALSE}
X <- matrix(c(rep(c(1,0),10), rep(c(0,1),10)), nrow = 20, byrow = TRUE)
colnames(X) <- c("T1", "T2")
se <- rep(1,20)
d <- c(5,20)

nsim <- 100
cols <- nsim
rows <- length(X[1,])
estTE <- matrix(NA, nrow = rows, ncol = cols)
cover_common <- 0
cover_random <- 0
taus <- c()
  
for (i in 1:nsim){
  simdata <- generate_2arm(X, d, se)
  net <- netmeta(TE, seTE, treat1, treat2, studlab, data = simdata)
  
  estTE[,i] <- net$TE.common[-1,1]  #TE at i-th simulation is stored in the i-th column
    
  cover_common <- cover_common + min(d <= net$upper.common[-1,1])* 
                                    min(d >= net$lower.common[-1,1])
  cover_random <- cover_random + min(d <= net$upper.random[-1,1])* 
                                    min(d >= net$lower.random[-1,1])
    
  taus <- c(taus, net$tau)
}
  
variance <- apply(estTE, 1, var) #take variance of each row
bias <- rowMeans(estTE-d)/d
MSE <- rowMeans((estTE-d)^2)
  
cat("Mean:", rowMeans(estTE), "\n")
cat("Variance:", variance, "\n")
cat("Relative bias:", bias, "\n")
cat("MSE:", MSE, "\n")
cat("Percentage coverage of CI (common):", cover_common/nsim, "\n")
cat("Percentage coverage of CI (random):", cover_random/nsim, "\n")
cat("Number of non-zero taus:", length(taus[which(taus != 0)]), "\n" )
cat("Mean of taus:", mean(taus), "\n")
```



\newpage
# simulate data of three models and a simple analysis
```{r}
set.seed(123)
```

Suppose we have treatments A, B, C, use A as the reference treatment, design matrix is given by 
```{r}
X <- matrix(c(rep(c(1,0),5), rep(c(0,1),5)), nrow = 10, byrow = TRUE)
colnames(X) <- c("B", "C")
X
```
standard error of each study is assumed to be 1, true treatment effects are assumed to be $\mu_{AB}=-2$, $\mu_{AC}=2$
```{r}
se <- rep(1,10)
d <- c(-2,2)
```

Consider three different models

## 1) Fixed effect model
$\phi=1, \tau=0$

use netmeta() for analysis
```{r}
phi1 <- 1
tau1 <- 0
data1 <- generate_2arm(X,mu,se,phi1,tau1,ref="A")
data1
```

```{r}
netmeta(TE, seTE, treat1, treat2, studlab, data = data1)
```

## 2) Mutiplicative effect
$\phi=2, \tau=0$
```{r}
phi2 <- 2
tau2 <- 0
data2 <- generate_2arm(X,mu,se,phi2,tau2,ref="A")
data2
```
weighted regression to estimate $\phi$
$$\theta = X\mu+\phi\cdot se^2$$
```{r}
fitme <- lm(data2$TE ~ X-1, weights = 1/(se^2))
summary(fitme)
```
```{r}
(summary(fitme)$sigma)^2
```

```{r}
netmeta(TE, seTE, treat1, treat2, studlab, data = data2)
```

## 3) Random effect model 
$\phi=1, \tau=1$
```{r}
phi3 <- 1
tau3 <- 1
data3 <- generate_2arm(X,mu,se,phi3,tau3,ref="A")
netmeta(TE, seTE, treat1, treat2, studlab, data = data3)
```



